%% Need from book [8] quotes: 
	%2 quotes for inferential level
	%2 quotes for appreciative level
	%2 quotes for evaluative level
	%2 quotes for critique level
%%AND
%Need from outside articles [2] quotes:
	%2 quotes for essential level(history|literary context)

Literal:
%Maya and Bailey are our two protagonists; they are young black children from LBC.

"Marguerite and Bailey Johnson Jr., from Long Beach, California, en route to Stamps, Arkansas, do Mrs. Annie Henderson" (Angelou 4).

%Maya's grandparents owned and ran a supply store in a plantation town in Arkansas.
"Then she had the Store built in the heart of the Negro area. Over the years it became the lay center of activities in town. On Saturdays, barbers sat their customers in the shade on the porch of the Store, and troubadours on their ceaseless crawlings through the South leaned across its benches and sang" (Angelou 4).

Crique (Tone): 
Self-loathing: Angelou will remind the reader that young Maya does not love herself because she was told by society that whiteness was equivalent to beauty.

Quote:
"I knew that once I put it on I'd look like a movie star. (It was silk and that made up for the awful color.) I was going to look like one of the sweet little white girls who were everybody's dream of what was right with the world" (Angelou 2).
"Wouldn't they be surprised when one day I woke out of my black ugly dream, and my real hair, which was long and blond, would take the place of the kinky mass that Momma wouldn't let me straighten? My light-blue eyes were going to hypnotize them, after all the things they said" (Angelou 3).


%%Unsorted quotes:

%Deep segregation in Arkansas
"In Stamps the segregation was so complete that most Black children didn't really, absolutely know what whites looked like. Other than that they were different, to be dreaded, and in that dread was included the hostility of the powerless against the powerful" (Angelou 13).

"These others, the strange pale creatures that lived in their alien unlife, weren't considered folks. They were whitefolks" (Angelou 13).


"Bailey said the man's things had been cut off and put in his pocket and he had been shot in the head, all because the whitefolks said he did 'it' to a white woman" (Angelou 18).


"If on Judgment Day I were summoned by St. Peter to give testimony to the used-to-be sheriff's act of kindness, I would be unable to say anything in his behalf. His confidence that my uncle and every other Black man who heard of the Klan's coming ride would scurry under their houses to hide in chicken droppings" (Angelou 10).

%Hope in the black community
"Whatever was given by Black people to other Blacks was most probably needed as desperately by the donor as by the receiver. A fact which made the giving or receiving a rich exchange" (Angelou 24).

%XMaya feels unworthy of her brother's love.
"She was his Mother Dear and I resigned myself to his condition. They were more alike than she and I, or even he and I. They both had physical beauty and personality, so I figured it figured" (Angelou 30).

%Distant feelings for her Californian father
"He was a stranger, and if he chose to leave us with a stranger, it was all of one piece" (Angelou 30).

"Occasionally Mother, whom we seldom saw in the house, had us meet her at Louie's. It was a long dark tavern at the end of the bridge near our school, and was owned by two Syrian brothers" (Angelou 32).


%Foreshadowing the recklessness of Mother, the danger of Mr. Freeman.
"Mr. Freeman moved gracefully, like a big brown bear, and seldom spoke to us. He simply waited for Mother and put his whole self into the waiting. He never read the paper or patted his foot to radio. He waited. That was all" (Angelou 35).


%The violent action of the rape of Maya.
"'If you scream, I'm gonna kill you. And if you tell, I'm gonna kill Bailey.' I could tell he meant what he said. I couldn't understand why he wanted to kill my brother. Neither of us had done anything to him. And then. Then there was the pain. A breaking and entering when even the senses are torn apart" (Angelou 38).


%After Maya's family takes justice into their own hands (police chief and organized crime side of her Mother's family)

"One lie surely wouldn't be worth a man's life. Bailey could have explained it all to me, but I didn't dare ask him. Obviously I had forfeited my place in heaven forever, and I was as gutless as the doll I had ripped to pieces ages ago" (Angelou 42).

%%Maya's maternal family get fed up with her PTSD or needs to cover their murdering tracks, and ships the siblings back to Arkansas.
%Maya feels dirty, as if she sinned:
%Evaluative: Maya’s self-hatred only grew worse following her rape, as she felt that she was broken,‘used goods.’ 
"[N]or did I want Uncle Willie, whom I loved in my fashion, to think of me as being sinful and dirty" (Angelou 45).

%%Prediction: Bailey comes back traumatized from town because he witnesses a lynching or other racist brutality.
"Why is he saying the baby prayer? We had been saying the "Our Father, which art in heaven" for years" (Angelou 57).


%%At a mega gathering of different denomination in town, the Maya feels the resentment:
"The mean whitefolks was going to get their comeuppance. Wasn't that what the minister said, and wasn't he quoting from the words of God Himself? They hadbeen refreshed with the hope of revenge and the promise of justice" (Angelou 63).

"A stranger to the music could not have made a distinction between the songs sung a few minutes before and those being danced to in the gay house by the railroad tracks. All asked the same questions. How long, oh God? How long?" (Angelou 63).

%%While listening to a black man in a boxing match against a white challenger for the championship. Maya's own racial value was on the line.
"[T]he accusations that we were lower types of human beings. Only a little higher than the apes. True that we were stupid and ugly and lazy and dirty and, unlucky and worst of all, that God Himself hated us" (Angelou 66).


%On her self-deprecating care for her new friend, Louise.
"Her hair was 'good'—more straight than kinky-—and her
features had the regularity of objects placed by a careful hand" (Angelou 69).

"Oh, it was important, all right. Whitefolks would attend the ceremony, and two or three would speak of God and home, and the Southern way of life ... IN the Store I was the person of the moment. The birthday girl" (Angelou 84).


%%Spite at Donleavy, the white, unplanned political speaker, who promised them breadcrumbs vs what he promised the white schools
"Donleavy had exposed us. We were maids and farmers, handymen and washerwomen, and anything higher that we aspired to was farcical and presumptuous. Then I wished that Gabriel Prosser and Nat Turner had killed all whitefolks in their beds and that Abraham Lincoln had been assassinated before the signing of the Emancipation Proclamation" (Angelou 84).


%%The racist Dentist Lincoln refused to help Maya as a returning favor to Mama (the grandmother).
"'Annie, my policy is I'd rather stick my hand in a dog's mouth than in a nigger's'" (Angelou 93).
